VM-Centric Deduplication with Fault Isolation for Snapshot Backup in the Cloud

### 2014-01-28 snapshot_system/paper_background.tex
update converged arch
Rich makes it through paper_dedup

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system (typically one of only a few
installed on the system) and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a seperate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup sevice may collocate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a seperate storage cluster.

@@

%As discussed earlier, collocating a backup service on the existing
%cloud cluster avoids the extra cost to acquire a dedicated backup facility

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In its simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
eficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

@@

Thus,
localizing deduplication to the VM can align fault isolation properties and
guarantees, improve the computational efficiency of snapshot maintenance, and
simplyfy snapshot deletion over non-localized approaches at the possible
expense of storage efficiency.

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block (FSB) size in a distributed file system such as  
the Hadoop file system and the Google file system
is uniform and large (e.g.  64MB) so as to optimize file system performance
for bandwidth usage.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distriuted among large file system blocks when a large-scale distributed file
system is used for disk image files.

@@

\paragraph*{Moderating Resource Usage} --
In a converged setting, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service must share the resources available to
other cloud services as illustrated in Figure~\ref{fig:collocated},
and thus miniomize its impact on user-deliverable performance.
%The key resource for global comparison  is memory for storing the fingerprints.
%During snapshot deltion, reference counting for all unique blocks used in all VMs requires a significant resource usage
%too.  Thus we seek for the low-profile techniques with approximation and low resource consumption.

>>>>

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system (typically one of only a few
installed on the system) and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

%As discussed earlier, co-locating a backup service on the existing
%cloud cluster avoids the extra cost to acquire a dedicated backup facility

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In its simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
efficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

@@

Thus,
localizing deduplication to the VM can align fault isolation properties and
guarantees, improve the computational efficiency of snapshot maintenance, and
simplify snapshot deletion over non-localized approaches at the possible
expense of storage efficiency.

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block (FSB) size in a distributed file system such as  
the Hadoop file system and the Google file system
is uniform and large (e.g.  64MB) so as to optimize file system performance
for bandwidth usage.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for disk image files.

@@

\paragraph*{Moderating Resource Usage} --
In a converged setting, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service must share the resources available to
other cloud services as illustrated in Figure~\ref{fig:collocated},
and thus minimize its impact on user-deliverable performance.
%The key resource for global comparison  is memory for storing the fingerprints.
%During snapshot deletion, reference counting for all unique blocks used in all VMs requires a significant resource usage
%too.  Thus we seek for the low-profile techniques with approximation and low resource consumption.

### 2014-01-28 snapshot_system/paper_dedup.tex
Rich makes it through paper_dedup

\subsection{VM centric  strategies}
\label{sect:vc-strategies}
%\begin{itemize}
%\item 
\textbf{VM-specific local duplicate search within similar segments.}
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.

@@

%\item 
\textbf{Global deduplication with popular chunks and replication support.}
This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is that the local deduplication has removed most of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
This number can be computed periodically, e.g., on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance.

@@

Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
We denote $\sigma$ as the percentage of unique data chunks beloning to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

%\item 
\textbf{VM-centric file system block management.}
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using a log-structured storage scheme built
on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

%factor on popular chunks. This explains why we must separate the PDS and
%non-PDS data in our system to acheive the desired $r_c/r$ ratio (though both
%can be stored on the same DFS).

>>>>

\subsection{VM centric  strategies}
\label{sect:vc-strategies}

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search optimization, global deduplication using popular chunks, and 
VM-centric file system block management. 
%\begin{itemize}
%\item 
\paragraph*{VM-local duplicate search optimization} --
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation 
%with Xen on an Alibaba platform, 
the segment size is 2MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.

@@

%\item 
\paragraph*{Global deduplication using popular chunks} --
This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is that the local deduplication has removed most of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
This number can be computed periodically, e.g., on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance.

@@

Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

%\item 
\paragraph*{VM-centric file system block management} --
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using a log-structured storage scheme built
on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

%factor on popular chunks. This explains why we must separate the PDS and
%non-PDS data in our system to achieve the desired $r_c/r$ ratio (though both
%can be stored on the same DFS).

### 2014-01-28 snapshot_system/paper_intro.tex
Rich makes it through paper_dedup

Without deduplication,
the cloud typically offloads backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degredation to production applications when 
backups are in progress.
However, sending out pre-deduplication backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).

In this paper we propose a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource swith them.
We describe methodologies to address the following challenges: 
\begin{itemize}
\item good global deduplication efficiency using a distributed implementation
that limits its resource usage, 
\item augmented fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception, and
\item efficient garbage collection so that users stop incurring 
storage charges immediately after snapshot deletion.
\end{itemize}

% low-cost VM-centric snapshot backup approach that is collocated within 
% the existing converged storage architecture. Our solution leverages the scalability of the underlying
% distributed file system and provides global data deduplication in a resource friendly manner.

@@

One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  The problem is particularly accute for VM
snapshots where the purpose of the snapshot is precisely to create redundant
data copies so that VMs are more fault resilient.  Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiencies that deduplication makes possible.

@@

This paper explores the storage and
performance efficiencies that such a {\em VM-centric} 
combination of local and global
techniques can achieve.  We term it VM-centric because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud enviornments.

%Because data
%sharing is restricted, this VM-centric approach reduces the overall resource
%usage significantly during backup and simplifies the snapshot deletion
%process. This low-resource design is suitable when the backup service with
%deduplication is collocated with other services running on a shared compute
%and storage cluster.  We have evaluated this VM-centric approach using  a
%prototype system.
% that runs a cluster of Linux machines with Xen and a standard distributed
% file system for the backup storage. 
%This approach localizes duplicate detection within each VM  By narrowing
%duplicate sharing within a small percent of common data chunks and exploiting
%their popularity, this scheme can afford to allocate extra replicas of these
%shared chunks for better fault resilience while sustaining competitive
%deduplication efficiency.

>>>>

Without deduplication,
the cloud typically offloads backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out pre-deduplication backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).

In this paper we propose a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource with them.
We describe methodologies to address the following challenges: 
\begin{itemize}
\item good global deduplication efficiency using a distributed implementation
that limits its resource usage, 
\item augmented fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception, and
\item efficient garbage collection so that users stop incurring 
storage charges immediately after snapshot deletion.
\end{itemize}

% low-cost VM-centric snapshot backup approach that is co-located within 
% the existing converged storage architecture. Our solution leverages the scalability of the underlying
% distributed file system and provides global data deduplication in a resource friendly manner.

@@

One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  The problem is particularly acute for VM
snapshots where the purpose of the snapshot is precisely to create redundant
data copies so that VMs are more fault resilient.  Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiencies that deduplication makes possible.

@@

This paper explores the storage and
performance efficiencies that such a {\em VM-centric} 
combination of local and global
techniques can achieve.  We term it VM-centric because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud environments.

%Because data
%sharing is restricted, this VM-centric approach reduces the overall resource
%usage significantly during backup and simplifies the snapshot deletion
%process. This low-resource design is suitable when the backup service with
%deduplication is co-located with other services running on a shared compute
%and storage cluster.  We have evaluated this VM-centric approach using  a
%prototype system.
% that runs a cluster of Linux machines with Xen and a standard distributed
% file system for the backup storage. 
%This approach localizes duplicate detection within each VM  By narrowing
%duplicate sharing within a small percent of common data chunks and exploiting
%their popularity, this scheme can afford to allocate extra replicas of these
%shared chunks for better fault resilience while sustaining competitive
%deduplication efficiency.
